---
title: "simpleRegress"
author: "Ashley Cook"
date: "July 5, 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```
  
If you look right above 20 that would hit neck, which is about 57 length wise. The line that shows across the chart, is the prediction. It is like a numerical variable, showing or explaining another numerical variable. The R console box shows us data that tells us how good the information is, more in depth. The residual standard error the smaller the number is, the bigger the model is. The one with the small residual error, will have have the biggest adgusted r squared. 

## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 


```{r}
MYCHEST <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(MYCHEST)
summary.lm(MYCHEST)
```
  
The residual error is 4.9 and the r squared is 0.791. The lower residual error is better. So, this chart, has a better residual error and data. The residual error that was on the other chart was 5.376, which is higher and not the better data. Chest is above neck, in rank, which means chest is better than neck. 

```{r}
MYHEADWTH <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(MYHEADWTH)
summary.lm(MYHEADWTH)
```

  
With this residual error, it has a higher value with a 7.3, which is not better than the other two charts that we had before this one. The chart is more spread out means the data is worse. The r squared is 0.5409, which is smaller than the other two r squared before. 


```{r}
MYWEIGHT <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(MYWEIGHT)
summary.lm(MYWEIGHT)
```

  
This residual standard error is a 5.432 which is a lower number which means the data is good and works better. The r squared number is 0.7472 which is low and is around the same as the other r squared numbers above. The line that goes across as known as the prediction, is going higher but is starting from a higher number also. All the dots are more closer together which means the data is works.

```{r}
MYAGE <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(MYAGE)
summary.lm(MYAGE)
```
  
The residual error is high with a 7.511 which is not good. While the r squared is at 0.5166, which is so far the best one out of all the plot charts we did. The smaller the r squared, the better outcome. 

```{r}
MYMONTH <- lm(LENGTH~MONTH,data=bears)
plot(LENGTH~MONTH,data=bears)
abline(MYMONTH)
summary.lm(MYMONTH)
```
  
This chart has the highest residual error of 10.99 which is bad. Then the r squared is 0.08423 which is now the lowest besides the AGE chart above. This data gives us a different chart, showing more blocks then dots. 

```{r}
MYSEX <- lm(LENGTH~SEX,data=bears)
plot(LENGTH~SEX,data=bears)
abline(MYSEX)
summary.lm(MYSEX)
```
  
The residual standard is 10.75 which is high and not good. The r squared is 0.01065 which is super good for this information for SEX. 
